---
title: "mtDNA"
output: html_document
date: "2024-10-29"
---
## Load packages and file locations etc.
```{r}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
library('gt') ; library('gtsummary')
```

## Load pilot data
```{r}
dat_mtDNA_Animnaming = readr::read_csv(paste0(pilot.folder,'mtDNA_Animnaming.csv'))
dat_mtDNA_calcvars = readr::read_csv(paste0(pilot.folder,'mtDNA_calcvars.csv'))
dat_mtDNA_climate_cvi = readr::read_csv(paste0(pilot.folder,'mtDNA_climate_cvi.csv'))
dat_mtDNA_Cog_impairment = readr::read_csv(paste0(pilot.folder,'mtDNA_Cog_impairment.csv'))
dat_mtDNA_Cog_Scores = readr::read_csv(paste0(pilot.folder,'mtDNA_Cog_Scores.csv'))
dat_mtDNA_Ids = readr::read_csv(paste0(pilot.folder,'mtDNA_Ids.csv'))
dat_mtDNA_letterF = readr::read_csv(paste0(pilot.folder,'mtDNA_letterF.csv'))
dat_mtDNA_moca = readr::read_csv(paste0(pilot.folder,'mtDNA_moca.csv'))
dat_mtDNA_sis = readr::read_csv(paste0(pilot.folder,'mtDNA_sis.csv'))
dat_mtDNA_WLL = readr::read_csv(paste0(pilot.folder,'mtDNA_WLL.csv'))
```

## Load processed mtDNA mutation data
```{r}
dat_mtDNA_mutations = readr::read_csv(paste0(pilot.folder,'variants_gene.csv'))
```

# Choose mutation to run model with as output
```{r}
seed_grid = names(dat_mtDNA_mutations)[c(16:ncol(dat_mtDNA_mutations))]
```

## Iterate over mutations and load model results
```{r}
model_0_results = data.frame(); model_1_results = data.frame() ; model_2_results = data.frame()
for(outcome_mutation in seed_grid){
  print(outcome_mutation)
  
  # Model 0: load csv
  model_0_results_current = read_csv(paste0(model.output.folder,"pilot_model0_",outcome_mutation,".csv")) %>%
    mutate(outcome=outcome_mutation)
  model_0_results = data.table::rbindlist(list(model_0_results,model_0_results_current))
  
  # Model 1: load csv
  model_1_results_current = read_csv(paste0(model.output.folder,"pilot_model1_",outcome_mutation,".csv")) %>%
    mutate(outcome=outcome_mutation)
  model_1_results = data.table::rbindlist(list(model_1_results,model_1_results_current))
  
  # Model 2: load csv
  model_2_results_current = read_csv(paste0(model.output.folder,"pilot_model2_",outcome_mutation,".csv")) %>%
    mutate(outcome=outcome_mutation)
  model_2_results = data.table::rbindlist(list(model_2_results,model_2_results_current))
}
```

## Save csv
```{r}
write_csv(model_0_results,paste0(model.output.folder,"pilot_model0_all_combined.csv"))
write_csv(model_1_results,paste0(model.output.folder,"pilot_model1_all_combined.csv"))
write_csv(model_2_results,paste0(model.output.folder,"pilot_model2_all_combined.csv"))
```


## Integrating the REGARDS datasets

# Install and load the moments package if not already installed
```{r}
library(moments)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggrepel)
```
#Reading all the datasets and merging them progressively
This current pilot analses do not include samples during follow-up
However, cognitive trajectories will be determined longitudinally
```{r}
#identifiers
ids <- read.csv("mtDNA_Ids.csv")
#covariates
vars <- read.csv("mtDNA_calcvars.csv")
#REGARDS dataset
regards <- read.csv("REGARDS-mtDNA-VariantsBySample.csv")
df_filtered <- regards %>%
  filter(!grepl("^rpt", SAMPLE))
df <- df_filtered %>%
  separate(SAMPLE, into = c("Sample_ID", "sample_2"), sep = "-")
```
#Merging IDs and covariates
```{r}
basic1 <- merge(ids, vars, by = "id", all = TRUE)
colnames(basic1)
basic1.2 <- merge(basic1, df, by = "Sample_ID")
summary(basic1.2$Wbc)
sd(basic1.2$Wbc, na.rm = TRUE)
table(basic1.2$Urbangrp)
```
#Descriptive statistics for all the covariates
Age
```{r}
colnames(basic1)
summary(basic1.2$Age)
sd(basic1.2$Age)
#basic histogram
hist(basic1$Age, 
     main = "Age of REGARDS participants", 
     xlab = "Age", 
     ylab = "Frequency", 
     col = "lightblue", 
     border = "black")
#Density plot 
plot(density(basic1$Age), 
     main = "Age of REGARDS participants", 
     xlab = "Age", 
     ylab = "Density", 
     col = "blue")
#Boxplot
boxplot(basic1$Age, 
        main = "Boxplot of Data", 
        ylab = "Values", 
        col = "lightgreen") #only one outlier
#Q-Q plot
qqnorm(basic1$Age)
qqline(basic1$Age, col = "red")
shapiro.test(basic1$Age) #it is not a normally distributed variable
#Skewness and kurtosis
# Calculate skewness and kurtosis
skewness(basic1$Age) #0.3775963 The distribution is positively skewed or right-skewed, meaning the right tail (higher values) is longer or fatter than the left tail.
kurtosis(basic1$Age) #2.573131 The distribution is platykurtic. It has lighter tails and a flatter peak than the normal distribution, indicating fewer data are in the tails and more are in the shoulders.
```
Race
```{r}
#colnames(basic1)
 table(basic1.2$Race)

#Race
freq_dist1 <- table(basic1$Race)
print(freq_dist1)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Race", 
        xlab = "Race", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Race", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Race)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Race", x = "Categories", y = "Absolute number of participants")
```
Gender
```{r}
table(basic1.2$Gender)

a <- table(basic1$Gender)
prop.table(a) #we have higher number of women in the subset

#Gender
freq_dist1 <- table(basic1$Gender)
print(freq_dist1)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Gender", 
        xlab = "Gender", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Gender", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Gender)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Gender", x = "Categories", y = "Absolute number of participants")
```
Income
```{r}
colnames(basic1)

table(basic1$Income)
prop.table(a) 
#Income=1; <5K
#Income=2; 5-10K 
#Income=3; 10-15K 
#Income=4; 15-20K 
#Income=5; 20-25K 
#Income=6; 25-35K 
#Income=7; 35-50K 
#Income=8; 50-75K 
#Income=9; 75-150K 
#Income=10; >150K 

#Income
freq_dist1 <- table(basic1$Income)
print(freq_dist1)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Income", 
        xlab = "Income", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Income", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Income)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Income", x = "Categories", y = "Absolute number of participants")
```
Income 4 categories
```{r}
colnames(basic1) #Income_4cat

table(basic1.2$Income_4cat)
prop.table(a) 
#Income_4cat='less than $20k';
#Income_4cat='$20k-$34k';
#Income_4cat='$35k-$74k';
#Income_4cat='$75k and above';

#Income_4cat, four categories
freq_dist1 <- table(basic1$Income_4cat)
print(freq_dist1)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Income (four categories)", 
        xlab = "Income", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Income", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Income_4cat)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Income (four categories)", x = "Categories", y = "Absolute number of participants")
```
Education 
ED_Cat
```{r}
table(basic1.2$ED_Cat) 

#ED_Cat='Less than high school';
#ED_Cat='High school graduate';
#ED_Cat='Some college';
#ED_Cat='College graduate and above';

#ED_cat, 
freq_dist1 <- table(basic1$ED_Cat)
print(freq_dist1)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Education", 
        xlab = "Physical Activity", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Education", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$ED_Cat)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Education", x = "Categories", y = "Absolute number of participants")
```
Physical activity
Exercise_cat
```{r}
colnames(basic1.2) 

#Exercise_cat='None';
#Exercise_cat='1 to 3 time per week';
#Exercise_cat='4 or more per week';

table(basic1.2$Exercise_cat)
print(freq_dist1)
#freq_dist2 <- table(basic1$Exercise_cat, useNA = "ifany")
#print(freq_dist2)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Physical Activity", 
        xlab = "Education", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Physical Activity", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Exercise_cat)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Physical Activity", x = "Categories", y = "Absolute number of participants")
```
Smoking
```{r}
colnames(basic1) 
#Smoke='Current';
#Smoke='Past';
#Smoke='Never';

table(basic1.2$Smoke)
print(freq_dist1)
#freq_dist2 <- table(basic1$Smoke, useNA = "ifany")
#print(freq_dist2)
prop_dist1 <- prop.table(freq_dist1)
print(prop_dist1)
summary_table <- data.frame(
  Category = names(freq_dist1),
  Frequency = as.vector(freq_dist1),
  Proportion = as.vector(prop_dist1)
)

# Bar plot using base R
barplot(freq_dist1, 
        main = "Bar Plot of Categories for Smoking", 
        xlab = "Education", 
        ylab = "Absolute number", 
        col = "lightblue", 
        border = "black")

# Pie chart using base R
pie(freq_dist1, 
    main = "Pie Chart of Categories for Smoking", 
    col = rainbow(length(freq_dist1)))

# Bar plot using ggplot2
# Create a data frame
data_df <- data.frame(Category = basic1$Smoke)
ggplot(data_df, aes(x = Category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Bar Plot of Categories for Smoking", x = "Categories", y = "Absolute number of participants")
```
BMI
```{r}
colnames(basic1) 

summary(basic1$BMI)
sd(basic1$BMI, )
#basic histogram
hist(basic1$BMI, 
     main = "BMI of REGARDS participants", 
     xlab = "BMI", 
     ylab = "Frequency", 
     col = "lightblue", 
     border = "black")
#Density plot 
# Remove NAs
#data_clean <- na.omit(basic1$BMI)
#plot(density(data_clean$BMI), 
#     main = "BMI of REGARDS participants", 
#     xlab = "BMI", 
#     ylab = "Density", 
#     col = "blue")
#Boxplot
boxplot(basic1$BMI, 
        main = "Boxplot of Data", 
        ylab = "Values", 
        col = "lightgreen") #only one outlier
#Q-Q plot
qqnorm(basic1$BMI)
qqline(basic1$BMI, col = "red")
shapiro.test(basic1$BMI) #it is not a normally distributed variable
#Skewness and kurtosis
# Calculate skewness and kurtosis
#skewness(basic1$BMI) 
#kurtosis(basic1$BMI) 
```


#############
## VOLCANOS PLOT




Mutation_count_effects_on_first_values_of_cognitive_measures
```{r}
a <- read.csv("AI5/Mutation_count_effects_on_first_values_of_cognitive_measures.csv")
first_z_mocam <- a %>% filter(Outcome == "first_z_mocam")
#Bonferroni
first_z_mocam$bonferroni <- p.adjust(first_z_mocam$p_value, method = "bonferroni")
first_z_mocam <- first_z_mocam %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
First z score MoCA
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.05)
#7.377759

first_z_mocam$sig <- "No"
first_z_mocam$sig[first_z_mocam$Coef > 0 & first_z_mocam$p_value <= 0.05] <- "Pos"
first_z_mocam$sig[first_z_mocam$Coef < 0 & first_z_mocam$p_value <= 0.05] <- "Pos"
first_z_mocam$sig[first_z_mocam$Coef < 0 & first_z_mocam$p_value > 0.05] <- "Neg"
first_z_mocam$sig[first_z_mocam$Coef > 0 & first_z_mocam$p_value > 0.05] <- "Neg"
table(first_z_mocam$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_z_mocam, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("First z score MoCA") +
  xlab("                                                   Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +  # Red line in p=0.05
  geom_hline(yintercept = -log10(0.05), col = "blue", linetype = 'dashed') + # Blue line in FDR=0.05
  coord_cartesian(ylim = c(0, max(-log10(first_z_mocam$p_value)) + 0.5), 
                  xlim = c(min(first_z_mocam$Coef) - 0.5, max(first_z_mocam$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Individual_mutation_effects_on_first_values_of_cognitive_measures
```{r, echo=FALSE}
b <- read.csv("Individual_mutation_effects_on_first_values_of_cognitive_measures.csv")
first_z_mocam2 <- b %>% filter(Outcome == "first_z_mocam")
#Bonferroni
first_z_mocam2$bonferroni <- p.adjust(first_z_mocam2$p_value, method = "bonferroni")
first_z_mocam2 <- first_z_mocam2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
First z score MoCA
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

first_z_mocam2$sig <- "No"
first_z_mocam2$sig[first_z_mocam2$Coef > 0 & first_z_mocam2$p_value <= 0.05] <- "Pos"
first_z_mocam2$sig[first_z_mocam2$Coef < 0 & first_z_mocam2$p_value <= 0.05] <- "Pos"
first_z_mocam2$sig[first_z_mocam2$Coef < 0 & first_z_mocam2$p_value > 0.05] <- "Neg"
first_z_mocam2$sig[first_z_mocam2$Coef > 0 & first_z_mocam2$p_value > 0.05] <- "Neg"
table(first_z_mocam2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_z_mocam2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) + # Increase the size of the points
  ggtitle("First z score MoCA") +
  xlab("                Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(first_z_mocam2$p_value)) + 0.5), 
                  xlim = c(min(first_z_mocam2$Coef) - 0.5, max(first_z_mocam2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
First z score for the overall composite with MoCA
```{r}
first_comp_overall2 <- b %>% filter(Outcome == "first_comp_overall2")
#Bonferroni
first_comp_overall2$bonferroni <- p.adjust(first_comp_overall2$p_value, method = "bonferroni")
first_comp_overall2 <- first_comp_overall2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")

#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

first_comp_overall2$sig <- "No"
first_comp_overall2$sig[first_comp_overall2$Coef > 0 & first_comp_overall2$p_value <= 0.05] <- "Pos"
first_comp_overall2$sig[first_comp_overall2$Coef < 0 & first_comp_overall2$p_value <= 0.05] <- "Pos"
first_comp_overall2$sig[first_comp_overall2$Coef < 0 & first_comp_overall2$p_value > 0.05] <- "Neg"
first_comp_overall2$sig[first_comp_overall2$Coef > 0 & first_comp_overall2$p_value > 0.05] <- "Neg"
table(first_comp_overall2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_comp_overall2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) + # Increase the size of the points
  ggtitle("First z score for the overall composite with MoCA") +
  xlab("                Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(first_comp_overall2$p_value)) + 0.5), 
                  xlim = c(min(first_comp_overall2$Coef) - 0.5, max(first_comp_overall2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Mutation_count_effects_on_raw_cognitive_values
```{r}
e <- read.csv("Mutation_count_effects_on_raw_cognitive_values.csv")
comp_fluency <- e %>% filter(Outcome == "comp_fluency")
#Bonferroni
comp_fluency$bonferroni <- p.adjust(comp_fluency$p_value, method = "bonferroni")
comp_fluency <- comp_fluency %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
Composite score for verbal fluency
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.05)
#7.377759

comp_fluency$sig <- "No"
comp_fluency$sig[comp_fluency$Coef > 0 & comp_fluency$p_value <= 0.05] <- "Pos"
comp_fluency$sig[comp_fluency$Coef < 0 & comp_fluency$p_value <= 0.05] <- "Pos"
comp_fluency$sig[comp_fluency$Coef < 0 & comp_fluency$p_value > 0.05] <- "Neg"
comp_fluency$sig[comp_fluency$Coef > 0 & comp_fluency$p_value > 0.05] <- "Neg"
table(comp_fluency$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = comp_fluency, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("Composite score for verbal fluency") +
  xlab("                       Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(comp_fluency$p_value)) + 0.5), 
                  xlim = c(min(comp_fluency$Coef) - 0.5, max(comp_fluency$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
z score for Animal fluency test
```{r}
z_af <- e %>% filter(Outcome == "z_af")
#Bonferroni
z_af$bonferroni <- p.adjust(z_af$p_value, method = "bonferroni")
z_af <- z_af %>% separate(Predictor, c("Origen", "mutation"), sep = "_")

# Threshold for significant mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)

z_af$sig <- "No"
z_af$sig[z_af$Coef > 0 & z_af$p_value <= 0.05] <- "Pos"
z_af$sig[z_af$Coef < 0 & z_af$p_value <= 0.05] <- "Pos"
z_af$sig[z_af$Coef < 0 & z_af$p_value > 0.05] <- "Neg"
z_af$sig[z_af$Coef > 0 & z_af$p_value > 0.05] <- "Neg"

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = z_af, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("z score for Animal fluency test") +
  xlab("                               Estimates") +
  ylab("-log10(p-value)") + # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(z_af$p_value)) + 0.5), 
                  xlim = c(min(z_af$Coef) - 0.5, max(z_af$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Individual_mutation_effects_on_raw_cognitive_values
```{r}
f <- read.csv("Individual_mutation_effects_on_raw_cognitive_values.csv")
comp_fluency_2 <- f %>% filter(Outcome == "comp_fluency")
#Bonferroni
comp_fluency_2$bonferroni <- p.adjust(comp_fluency_2$p_value, method = "bonferroni")
comp_fluency_2 <- comp_fluency_2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
Composite score for verbal fluency
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

comp_fluency_2$sig <- "No"
comp_fluency_2$sig[comp_fluency_2$Coef > 0 & comp_fluency_2$p_value <= 0.05] <- "Pos"
comp_fluency_2$sig[comp_fluency_2$Coef < 0 & comp_fluency_2$p_value <= 0.05] <- "Pos"
comp_fluency_2$sig[comp_fluency_2$Coef < 0 & comp_fluency_2$p_value > 0.05] <- "Neg"
comp_fluency_2$sig[comp_fluency_2$Coef > 0 & comp_fluency_2$p_value > 0.05] <- "Neg"
table(comp_fluency_2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = comp_fluency_2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("Composite score for verbal fluency") +
  xlab("                               Estimates") +
  ylab("-log10(p-value)") + # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(comp_fluency_2$p_value)) + 0.5), 
                  xlim = c(min(comp_fluency_2$Coef) - 0.5, max(comp_fluency_2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```












