# Install and load the moments package if not already installed
```{r}
library(moments)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggrepel)
```
#############
## VOLCANOS PLOT




Mutation_count_effects_on_first_values_of_cognitive_measures
```{r}
a <- read.csv("AI5/Mutation_count_effects_on_first_values_of_cognitive_measures.csv")
first_z_mocam <- a %>% filter(Outcome == "first_z_mocam")
#Bonferroni
first_z_mocam$bonferroni <- p.adjust(first_z_mocam$p_value, method = "bonferroni")
first_z_mocam <- first_z_mocam %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
First z score MoCA
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.05)
#7.377759

first_z_mocam$sig <- "No"
first_z_mocam$sig[first_z_mocam$Coef > 0 & first_z_mocam$p_value <= 0.05] <- "Pos"
first_z_mocam$sig[first_z_mocam$Coef < 0 & first_z_mocam$p_value <= 0.05] <- "Pos"
first_z_mocam$sig[first_z_mocam$Coef < 0 & first_z_mocam$p_value > 0.05] <- "Neg"
first_z_mocam$sig[first_z_mocam$Coef > 0 & first_z_mocam$p_value > 0.05] <- "Neg"
table(first_z_mocam$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_z_mocam, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("First z score MoCA") +
  xlab("                                                   Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +  # Red line in p=0.05
  geom_hline(yintercept = -log10(0.05), col = "blue", linetype = 'dashed') + # Blue line in FDR=0.05
  coord_cartesian(ylim = c(0, max(-log10(first_z_mocam$p_value)) + 0.5), 
                  xlim = c(min(first_z_mocam$Coef) - 0.5, max(first_z_mocam$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Individual_mutation_effects_on_first_values_of_cognitive_measures
```{r, echo=FALSE}
b <- read.csv("Individual_mutation_effects_on_first_values_of_cognitive_measures.csv")
first_z_mocam2 <- b %>% filter(Outcome == "first_z_mocam")
#Bonferroni
first_z_mocam2$bonferroni <- p.adjust(first_z_mocam2$p_value, method = "bonferroni")
first_z_mocam2 <- first_z_mocam2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
First z score MoCA
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

first_z_mocam2$sig <- "No"
first_z_mocam2$sig[first_z_mocam2$Coef > 0 & first_z_mocam2$p_value <= 0.05] <- "Pos"
first_z_mocam2$sig[first_z_mocam2$Coef < 0 & first_z_mocam2$p_value <= 0.05] <- "Pos"
first_z_mocam2$sig[first_z_mocam2$Coef < 0 & first_z_mocam2$p_value > 0.05] <- "Neg"
first_z_mocam2$sig[first_z_mocam2$Coef > 0 & first_z_mocam2$p_value > 0.05] <- "Neg"
table(first_z_mocam2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_z_mocam2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) + # Increase the size of the points
  ggtitle("First z score MoCA") +
  xlab("                Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(first_z_mocam2$p_value)) + 0.5), 
                  xlim = c(min(first_z_mocam2$Coef) - 0.5, max(first_z_mocam2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
First z score for the overall composite with MoCA
```{r}
first_comp_overall2 <- b %>% filter(Outcome == "first_comp_overall2")
#Bonferroni
first_comp_overall2$bonferroni <- p.adjust(first_comp_overall2$p_value, method = "bonferroni")
first_comp_overall2 <- first_comp_overall2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")

#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

first_comp_overall2$sig <- "No"
first_comp_overall2$sig[first_comp_overall2$Coef > 0 & first_comp_overall2$p_value <= 0.05] <- "Pos"
first_comp_overall2$sig[first_comp_overall2$Coef < 0 & first_comp_overall2$p_value <= 0.05] <- "Pos"
first_comp_overall2$sig[first_comp_overall2$Coef < 0 & first_comp_overall2$p_value > 0.05] <- "Neg"
first_comp_overall2$sig[first_comp_overall2$Coef > 0 & first_comp_overall2$p_value > 0.05] <- "Neg"
table(first_comp_overall2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = first_comp_overall2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) + # Increase the size of the points
  ggtitle("First z score for the overall composite with MoCA") +
  xlab("                Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(first_comp_overall2$p_value)) + 0.5), 
                  xlim = c(min(first_comp_overall2$Coef) - 0.5, max(first_comp_overall2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Mutation_count_effects_on_raw_cognitive_values
```{r}
e <- read.csv("Mutation_count_effects_on_raw_cognitive_values.csv")
comp_fluency <- e %>% filter(Outcome == "comp_fluency")
#Bonferroni
comp_fluency$bonferroni <- p.adjust(comp_fluency$p_value, method = "bonferroni")
comp_fluency <- comp_fluency %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
Composite score for verbal fluency
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.05)
#7.377759

comp_fluency$sig <- "No"
comp_fluency$sig[comp_fluency$Coef > 0 & comp_fluency$p_value <= 0.05] <- "Pos"
comp_fluency$sig[comp_fluency$Coef < 0 & comp_fluency$p_value <= 0.05] <- "Pos"
comp_fluency$sig[comp_fluency$Coef < 0 & comp_fluency$p_value > 0.05] <- "Neg"
comp_fluency$sig[comp_fluency$Coef > 0 & comp_fluency$p_value > 0.05] <- "Neg"
table(comp_fluency$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = comp_fluency, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("Composite score for verbal fluency") +
  xlab("                       Estimates") +
  ylab("-log10(p-value)") +  # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(comp_fluency$p_value)) + 0.5), 
                  xlim = c(min(comp_fluency$Coef) - 0.5, max(comp_fluency$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
z score for Animal fluency test
```{r}
z_af <- e %>% filter(Outcome == "z_af")
#Bonferroni
z_af$bonferroni <- p.adjust(z_af$p_value, method = "bonferroni")
z_af <- z_af %>% separate(Predictor, c("Origen", "mutation"), sep = "_")

# Threshold for significant mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)

z_af$sig <- "No"
z_af$sig[z_af$Coef > 0 & z_af$p_value <= 0.05] <- "Pos"
z_af$sig[z_af$Coef < 0 & z_af$p_value <= 0.05] <- "Pos"
z_af$sig[z_af$Coef < 0 & z_af$p_value > 0.05] <- "Neg"
z_af$sig[z_af$Coef > 0 & z_af$p_value > 0.05] <- "Neg"

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = z_af, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("z score for Animal fluency test") +
  xlab("                               Estimates") +
  ylab("-log10(p-value)") + # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(z_af$p_value)) + 0.5), 
                  xlim = c(min(z_af$Coef) - 0.5, max(z_af$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
Individual_mutation_effects_on_raw_cognitive_values
```{r}
f <- read.csv("Individual_mutation_effects_on_raw_cognitive_values.csv")
comp_fluency_2 <- f %>% filter(Outcome == "comp_fluency")
#Bonferroni
comp_fluency_2$bonferroni <- p.adjust(comp_fluency_2$p_value, method = "bonferroni")
comp_fluency_2 <- comp_fluency_2 %>% separate(Predictor, c("Origen", "mutation"), sep = "_")
```
Composite score for verbal fluency
```{r}
#threshold for significan mutations at the Bonferroni threshold (0.05/8 mutations= 0.00625)
fdr_threshold <- -log10(0.2)
#7.377759

comp_fluency_2$sig <- "No"
comp_fluency_2$sig[comp_fluency_2$Coef > 0 & comp_fluency_2$p_value <= 0.05] <- "Pos"
comp_fluency_2$sig[comp_fluency_2$Coef < 0 & comp_fluency_2$p_value <= 0.05] <- "Pos"
comp_fluency_2$sig[comp_fluency_2$Coef < 0 & comp_fluency_2$p_value > 0.05] <- "Neg"
comp_fluency_2$sig[comp_fluency_2$Coef > 0 & comp_fluency_2$p_value > 0.05] <- "Neg"
table(comp_fluency_2$sig, useNA = c("ifany"))

#Apply a theme globally with larger sizes
theme_set(theme_minimal(base_size = 16))

#Create the Volcano
vulcano_plot <- ggplot(data = comp_fluency_2, aes(x = Coef, y = -log10(p_value), col = sig)) +
  geom_point(size = 2) +  # Increase the size of the points
  ggtitle("Composite score for verbal fluency") +
  xlab("                               Estimates") +
  ylab("-log10(p-value)") + # Modify Y-axis text to remove underscore
  geom_vline(xintercept = 0, col = "black", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "red", linetype = 'dashed') +
  geom_hline(yintercept = fdr_threshold, col = "blue", linetype = 'dashed') +
  coord_cartesian(ylim = c(0, max(-log10(comp_fluency_2$p_value)) + 0.5), 
                  xlim = c(min(comp_fluency_2$Coef) - 0.5, max(comp_fluency_2$Coef) + 0.5)) +
  scale_color_manual(values = c("Pos" = "blue", "Neg" = "red", "No" = "gray")) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5), # Increase title size
    axis.title = element_text(size = 14),               # Increase size of axis titles
    axis.text = element_text(size = 14)                 # Increase size of axis numbers
  )

# Create the chart with labels and white background
vulcano_plot_final <- vulcano_plot + 
  geom_text_repel(
    aes(label = Origen),
    size = 5,  # Increase the size of labels
    box.padding = 0.5,
    max.overlaps = 10
  )

#Show final graph
vulcano_plot_final
```
